---
title: "microbial visualization"
output: html_document
---

paper: https://journals.asm.org/doi/epub/10.1128/AEM.00149-21

This document shows some reproductions of the visualization in the paper above as well as some plots that give an overall understanding about the 3 datasets. (The graphs look better if knitting the document) Also, it includes some hypothesis tests for analyzing the effects of different microbial fungicides on the dollar spot.

```{r read data, message=FALSE}
soil = read.csv("./data/ASV_soil_norm_filt.txt", sep = "\t")
taxa = read.csv("./data/taxaonomy_variation_filt.txt", sep = "\t")
vari = read.csv("./data/Variation_Metadata_adj_2group.txt", sep = "\t")
```

```{r load packages, message=FALSE}
library(tidyverse)
library(Polychrome)
library(cowplot)
library(dplyr)
library(glasso)
library(BigQuic)
library(corrplot)
#library(colorRamps);library(igraph);library(RColorBrewer);library(threejs);library(htmlwidgets);
```

```{r DAI, message=FALSE}
# a dataset consists of columns of days after inoculation, percentage of greeness, sample number, highly/moderately susceptible
dai = data.frame(day = c(rep(0, 18), rep(2, 18), rep(4, 18), 
                    rep(6, 18), rep(8, 18), rep(10, 18), 
                    rep(12, 18), rep(14, 18), rep(16, 18)),
            greeness = c(vari$Severity0_Raw, vari$Severity2_Raw,
            vari$Severity4_Raw, vari$Severity6_Raw,
            vari$Severity8_Raw, vari$Severity10_Raw,
            vari$Severity12_Raw, vari$Severity14_Raw,
            vari$Severity16_Raw),
            category = rep(c("sample 1", "sample 2", "sample 3",
                             "sample 4", "sample 5", "sample 6",
                             "sample 7", "sample 8", "sample 9",
                             "sample 10", "sample 11", 
                             "sample 12", "sample 13", 
                             "sample 14", "sample 15", 
                             "sample 16", "sample 17", 
                             "sample 18"), 9),
            hs = rep(vari$SeverityPeak == "HS", 9),
            ms = rep(vari$SeverityPeak == "MS", 9))

# a curve plot about the changing percentage of the greeness in each sample
ggplot(data = dai, mapping = aes(x = day, y = greeness)) +
  geom_line(aes(colour = category)) +
  ggtitle("Dollar spot development as indicated by turf greeness decay")

# Dollar spot development as indicated by turf greenness decay curve fitted with sigmoidal model throughout 16 days of incubation after dollar spot inoculation
x = dai$day
y = dai$greeness
plot((dai[dai$hs == TRUE,])$greeness ~ (dai[dai$hs == TRUE,])$day, xlab = "Days after inoculation", ylab = "% Greeness", main = "The sigmoidal model of the percentage of greeness", col = "blue")
points((dai[dai$hs != TRUE,])$greeness ~ (dai[dai$hs != TRUE,])$day)
fit = nls(y ~ SSlogis(x, Asym, xmid, scal), data = data.frame(x, y))
lines(seq(0, 16, length.out = 100), 
      predict(fit, newdata = data.frame(x = seq(0, 16, length.out = 100))), col = "red")
legend("bottomleft", 
       legend = c("MS", "HS", "Sigmoidal Fitting"),
       col = c("black", "blue", "red"), pch = c(1, 1, NA),
       lty = c(0,0,1), cex=0.8)

# comparison between HS and MS curve plots
ggplot(data = dai,
       mapping = aes(x = day, y = greeness)) +
  geom_point(aes(colour = hs)) +
  geom_smooth(aes(color = hs), se = FALSE)+
  ggtitle("Comparison between HS and MS")

# Side-by-side boxplot about the average % turfgrass greenness over the peak disease development stage (4-10 days after inoculation) for both highly susceptible and moderately susceptible samples
ggplot(vari[vari$SeverityPeak != "Med", ], aes(factor(SeverityPeak), Peak_AVG_raw, color = SeverityPeak)) +
  geom_boxplot() + 
  geom_point() +
  ggtitle("Greenness over the peak disease development stage") +
  xlab("")

# t-test for the difference between HS and MS
# The small p-value indicates there is difference, rejecting H0
fit_ms = lm(greeness~day, data = dai[dai$ms == TRUE,])
fit_hs = lm(greeness~day, data = dai[dai$hs == TRUE,])
t.test((dai[dai$ms == TRUE,])$greeness, 
       (dai[dai$hs == TRUE,])$greeness)
```

```{r taxaonomy, message=FALSE, warning=FALSE}
# The dataset consists of the relative abundance of microbiome family for the highly susceptible samples
df_hs = data.frame(cat = unique(as.factor(taxa$Family)),
                count = 0)

for (i in 1:10992)
  df_hs[df_hs$cat == taxa[i,6],2] = 
  df_hs[df_hs$cat == taxa[i,6],2] + sum((soil[,c(TRUE, vari$SeverityPeak == "HS")])[i,-1])
df_hs$count = df_hs$count/sum(df_hs$count)

# The dataset consists of the relative abundance of microbiome family for the moderately susceptible samples
df_ms = data.frame(cat = unique(as.factor(taxa$Family)),
                count = 0)
for (i in 1:10992)
  df_ms[df_ms$cat == taxa[i,6],2]  = 
  df_ms[df_ms$cat == taxa[i,6],2] + sum((soil[,c(TRUE, vari$SeverityPeak == "MS")])[i,-1])
df_ms$count = df_ms$count/sum(df_ms$count)

# The dataset binds df_ms and df_hs together
family = rbind(df_ms, df_hs)
family = cbind(family, group = c(rep("MS", 269), rep("HS", 269)))

#Average relative abundances of rhizosphere microbiome from MS and HS turfgrass for the taxa that represent more than 1% of the identified community at family level

#!!MODIFICATION: try to make color discrete
ggplot(data = family[family$count > 0.01,], mapping = aes(x = group, y = count, color = cat)) +
  geom_col(aes(fill = cat), colour= "black", position = "fill",
           alpha = 0.5)+
  ggtitle("Relative abundance of microbiome(over 1% of the identified community at family level)") +
  theme_classic()
```

```{r hypothesis test, message=FALSE, warning=FALSE}
# In this dataset, the first column is the relative frequency of each family in HS, and the second column is that in MS
df = bind_cols(df_hs, df_ms)
df = df[,-c(1,3)]
df = sapply(df, as.numeric)
# using t test to check if there are differences in the relative abundance of family types between HS and MS samples 
t.test(df[,1], df[,2], alternative = "two.sided",
       paired = TRUE)
# delete the two unidentified families
t.test(df[-c(1, 2), 1], df[-c(1, 2), 2], alternative = "two.sided", 
       paired = TRUE)

# Conduct rank-sum test (not normally distributed) on the relative frequency of each family in HS and MS
p = c()
sample = as.data.frame(x = c(rep(0, 18)))
for (i in 1:length(unique(taxa$Family))){
  x = colSums(soil[taxa$Family== (unique(taxa$Family))[i], -1])
  sample = cbind(sample, x)
  p = c(p, wilcox.test(x[vari$SeverityPeak == "HS"], 
       x[vari$SeverityPeak != "HS"])$p.value)
}
sample = sample[,-1]
colnames(sample) = unique(taxa$Family)
rst = bind_cols(family = unique(taxa$Family), p_value = p)
head(rst, 10)
# how many of the relative frequency of family in HS and MS are different
sum((rst$p_value)[!is.na(rst$p_value)] > 0.05)
```

```{r difference of top family, message=FALSE}
df=bind_cols(df_hs, df_ms)
colnames(df) = c("cat_hs", "count_hs", "cat_ms", "count_ms")
df = df%>%
  mutate(diff = abs(count_hs - count_ms))
df = df[order(-df$diff), ]

a = t(df[1:10,-c(1, 3, 5)])
colnames(a) = df[1:10, 1]
barplot(a, beside = TRUE, cex.names = 0.4, 
        main = "The top 10 differences of relative frequency of family")
```

```{r correlation matrix, message=FALSE}
for (i in 1:18)
  sample[i,] = sample[i,]/sample[i,1]

x = (cor(sample))[-1, -1]

sig = 0.5
  #convert data to numeric in order to run correlations
  #convert to factor first to keep the integrity of the data - each value will become a number rather than turn into NA
  df_cor = sample %>% mutate_if(is.character, as.factor)
  df_cor = df_cor %>% mutate_if(is.factor, as.numeric)
  #run a correlation and drop the insignificant ones
  corr = cor(df_cor)
  #prepare to drop duplicates and correlations of 1     
  corr[lower.tri(corr,diag=TRUE)] = NA 
  #drop perfect correlations
  corr[corr == 1] = NA 
  #turn into a 3-column table
  corr = as.data.frame(as.table(corr))
  #remove the NA values from above 
  corr = na.omit(corr) 
  #select significant values  
  corr = subset(corr, abs(Freq) > sig) 
  #sort by highest correlation
  corr = corr[order(-abs(corr$Freq)),] 
  #print table
  head(corr, 10)
  #turn corr back into matrix in order to plot with corrplot
  mtx_corr = reshape2::acast(corr, Var1~Var2, value.var="Freq")
  #plot correlations visually
  #corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ", type = 'lower', diag = FALSE)
  
#heatmap(solve(x[1:40, 1:40]))
  
```


rho = 0.75 #regularization parameter for lasso 
invcov = glasso(x, rho = rho)
P = invcov$wi #the sparse inverse covariance matrix (conditional distribution)
colnames(P) = colnames(x)
rownames(P) = rownames(x)
# check symmetry
if (!isSymmetric(P)) {
  P[lower.tri(P)] = t(P)[lower.tri(P)]  
}
# calculate partial correlation matrix
parr.corr = matrix(nrow=nrow(P), ncol=ncol(P))
for(k in 1:nrow(parr.corr)) {
  for(j in 1:ncol(parr.corr)) {
    parr.corr[j, k] = -P[j,k]/sqrt(P[j,j]*P[k,k])
  }
}
colnames(parr.corr) = colnames(P)
rownames(parr.corr) = colnames(P)
diag(parr.corr) = 0

# build network graph
graph = graph_from_adjacency_matrix(parr.corr,
                                    mode="undirected",
                                    weighted=TRUE)
# drop vertices with no edges
isolated =  which(degree(graph) == 0)
graph = delete.vertices(graph, isolated)
# make interactive graph
graph_js = graphjs(g=graph, 
                   layout_with_fr(graph, 
                  weights = 30 * E(graph)$width, dim=3),  #`?layout` to get a list
                # vertex.shape = names(V(ig_wt)),  # plot nodes as tickers rather than circles
                vertex.size=0.7,
                vertex.frame.color="white",
                vertex.frame.width=0.2,
                vertex.label=names(V(graph)),  
                # label nodes with tickers
                brush=TRUE,  
                # enable highlighting clicked nodes and their connections
                showLabels=TRUE,  
                # show node labels on hover
                edge.alpha=0.6,  
                # edge opacity - lower helps when there are dense connections
                bg="white",  # background colour
                main="Network graph from Graphical Lasso")
# save graph
graph_filename = paste0("./network_graph_rho_", rho, ".html")
saveWidget(graph_js, file=graph_filename)
# open in browser
browseURL(graph_filename)



```{r big-quic}
ld = seq(from = 0.9, to = 0.99, by = 0.01)
quic = BigQuic(data.matrix(sample[, -1]), lambda = ld,memory_size = 512, seed = 1,
use_ram = TRUE)
# select the optimal lamda value from it
quic_pre = BigQuic.select(quic)
quic_pre = quic_pre$precision_matrices
a = as.matrix(quic_pre[[1]])
# negative relation: green
# positive relation: red
# symmetric
plot(quic)
```


