---
title: "Network Analysis"
output: html_document
---

This document uses glasso, correlation matrix, and dissimilarity matrix to construct microbial networks.

```{r load, message=FALSE}
library(glasso)
library(qgraph)
library(igraph)
library(threejs)
library(htmlwidgets)
library(vegan)
library(MCL)
sample = read.csv("sample.csv")
row.names(sample) = sample[,1]
sample = sample[,-c(1)]
```

```{r dissimilarity}
# Create dissimilarity matrix
diss.mat = as.matrix(vegdist(t(sample[,-1]), method = "bray"))
diss.cutoff = 0.1
diss.adj = ifelse(diss.mat <= diss.cutoff, 1, 0)
# Construct microbiome network from adjacency matrix
diss.net = graph.adjacency(diss.adj, mode = "undirected",
                           diag = FALSE)
```

```{r correlation}
cor.matrix = cor(sample[,-1], method = "pearson")
# Convert correlation matrix to binary adjacency matrix
cor.cutoff = 0.75
cor.adj = ifelse(abs(cor.matrix) >= cor.cutoff, 1, 0)
# Construct microbiome network from adjacency matrix
cor.net = graph.adjacency(cor.adj,
mode = "undirected",
diag = FALSE)
```

```{r glasso}
x = (cor(sample))[-1, -1]
rho = 0.75 #regularization parameter for lasso 
invcov = glasso(x, rho = rho)
P = invcov$wi #the sparse inverse covariance matrix (conditional distribution)
colnames(P) = colnames(x)
rownames(P) = rownames(x)
# check symmetry
if (!isSymmetric(P)) {
  P[lower.tri(P)] = t(P)[lower.tri(P)]  
}
# calculate partial correlation matrix
parr.corr = matrix(nrow=nrow(P), ncol=ncol(P))
for(k in 1:nrow(parr.corr)) {
  for(j in 1:ncol(parr.corr)) {
    parr.corr[j, k] = -P[j,k]/sqrt(P[j,j]*P[k,k])
  }
}
colnames(parr.corr) = colnames(P)
rownames(parr.corr) = colnames(P)
diag(parr.corr) = 0

# build network graph
glasso.net = graph_from_adjacency_matrix(parr.corr,
                                    mode="undirected",
                                    weighted=TRUE)
# drop vertices with no edges
# isolated =  which(degree(glasso.net) == 0)
# glasso.net = delete.vertices(glasso.net, isolated)
```

# make interactive graph
graph_js = graphjs(g=graph, 
                   layout_with_fr(graph, 
                  weights = 30 * E(graph)$width, dim=3),  #`?layout` to get a list
                # vertex.shape = names(V(ig_wt)),  # plot nodes as tickers rather than circles
                vertex.size=0.7,
                vertex.frame.color="white",
                vertex.frame.width=0.2,
                vertex.label=names(V(graph)),  
                # label nodes with tickers
                brush=TRUE,  
                # enable highlighting clicked nodes and their connections
                showLabels=TRUE,  
                # show node labels on hover
                edge.alpha=0.6,  
                # edge opacity - lower helps when there are dense connections
                bg="white",  # background colour
                main="Network graph from Graphical Lasso")
# save graph
graph_filename = paste0("./network_graph_rho_", rho, ".html")
saveWidget(graph_js, file=graph_filename)
# open in browser
browseURL(graph_filename)


```{r analysis}
net = glasso.net
#net = diss.net
#net = cor.net

# convert negative edge weight to positive
E(net)$weight = E(net)$weight - min(E(net)$weight)


#E(net)$weight = as.numeric(floor(E(net)$weight*10^7))
#https://github.com/igraph/igraph/issues/1009
#net.cn = closeness(net, mode = "all")
#net.bn = betweenness(net)
#net.pr = page_rank(net)$vector
# hubs: highly connected nodes in a network
net.hs = hub_score(net)$vector

# plot of abundance vs. centrality
plot(x = c(sample[1,-1]), y = c(net.hs), 
     xlab = "Relative abundance", ylab = "Hubbiness score")
#low abundance, high centrality
head(sort(net.hs, decreasing=TRUE), n = 9)


# Get clusters
# short random walks tend to stay in the same community
wt = walktrap.community(net)
#ml = multilevel.community(net)
# Get membership of walktrap clusters
# membership(wt)
# Get clusters using MCL method
#adj = as_adjacency_matrix(net)
#mc = mcl(adj, addLoops = TRUE)

plot_dendrogram(wt)
centralityPlot(net, 
               include = c("Degree", "Closeness","Betweenness"),
               orderBy = "Degree")

deg = degree(net, mode = "all")
# Degree distribution
deg.dist = degree_distribution(net, mode = "all", cumulative = T)
# Plot degree distribution
plot(deg.dist, xlab = "Nodes degree", ylab = "Probability")
lines(deg.dist)

#(removing it disconnect the graph)
AP = articulation.points(net)

# get the membership of the largest cluster
group = sapply(groups(wt), length)
key = wt[group == max(group)]
paste( unlist(key), collapse=', ')


# Function that plots network with clusters and node size scaled to hubbiness.Articulation points are circled. Different colors of the nodes represent different communities. All isolated nodes are white.
# scores: hubbiness scores
# cls: cluster walkstrao
# AP: articulation points
plot.net.cls = function(net, scores, cls, AP, outfile, title) {
  # Get size of clusters to find isolated nodes.
  # groups: get a list of vertex groups from community object
  cls_sizes = sapply(groups(cls), length)
  # Randomly choosing node colors
  #colors = sample(colours(), length(cls))
  colors = rainbow(length(cls))
  
  # Nodes in clusters will be color coded
  # Isolated nodes will be white.
  V(net)$color = sapply(membership(cls),
                         function(x) {
                            ifelse(cls_sizes[x]>1, 
                                   colors[x],
                                   "white")
                            }
                          )
  # Convert node label from names to numerical IDs.
  node.names = V(net)$name
  col_ids = seq(1, length(node.names))
  V(net)$name = col_ids
  # To draw a halo around articulation points.
  AP = lapply(names(AP), function(x) x)
  marks = lapply(1:length(AP), 
                  function(x) which(node.names == AP[[x]]))
  # Define output image file.
  outfile = paste(outfile, "jpg", sep=".")
  # Image properties.
  jpeg(outfile, width = 5800, height = 14000, res = 300, 
       quality = 100)
  par(oma = c(4, 1, 1, 1))
  # Customized layout to avoid nodes overlapping.
  e = get.edgelist(net)
  class(e) = "numeric"
  #get the layout of the graph
  l = qgraph.layout.fruchtermanreingold(e,
    # area default is the square of number of nodes
    vcount=vcount(net), area=8*(vcount(net)^2))
  # Main plot function.
  plot(net, vertex.size = (scores*5)+4, vertex.label.cex=0.9,
       vertex.label.color = "black", mark.border="black",
       mark.groups = marks, mark.col = "white", 
       mark.expand = 10, mark.shape = 1, layout=l)
  title(title, cex.main=4)
  # Plot legend containing taxa names
  labels = paste(as.character(V(net)), node.names, sep = ") ")
  legend("bottom", legend = labels, xpd = TRUE, ncol = 5, 
         cex = 1.2)
  dev.off()
}
plot.net.cls(net, net.hs, wt, AP, outfile = "network_graph",
title = "Microbial Network")
```

